---
title: "Data Wrangling Exercise 1 v2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting started

The data is in an Excel file [here](https://drive.google.com/file/d/0B9vKjeWdQHa5OE5UQ1k4TWJlelU/view) called refine.xlsx. Right away, you'll notice that the data set has a few issues:

There are four brands: Philips, Akzo, Van Houten and Unilever. However, there are many different spellings and capitalizations of those names!

The product code and number are combined in one column, separated by a hyphen. 

Exercise

Using R, clean this data set to make it easier to visualize and analyze. Specifically, these are the tasks you need to do: 

```{r refine}

```

### 0: Load the data in RStudio

Save the data set as a CSV file called refine_original.csv and load it in RStudio into a data frame.

  + Downloaded xlsx file from link
  + Opened R studio and import dataset as xls (no access to xls to convert to CSV format)
  + Installed tidyr, dplyr, knitr, readxls packages

``` {r eval = TRUE, message = FALSE}
library(tidyr)
library(dplyr)
library(knitr)
library(readxl)
refine <- read_excel("Source/refine.xlsx")
refine
```

### 1: Clean up brand names

Clean up the 'company' column so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

  + Selected the company column and turned it into a vector of names
  + Sorted the list of company names in alpha order to make creating a new dataframe easier
  + Used the vector to create a dataframe that corresponds to the correct spelling
  + Used left_join() to create new column with correct company name
  + Modified columns by using the select() function to get the columns in order
  + Renamed new column name back to "company"

``` {R eval = TRUE}
sort_company <- sort(as.vector(as.matrix(refine[,1])))
sort_company_df <- data_frame(company = sort_company, fixed_company = c("akzo",  "akzo",  "akzo",  "akzo",  "akzo",  "akzo",  "akzo", "philips", "philips", "philips", "philips", "philips", "philips", "philips", "philips", "philips", "van houten",  "van houten",  "van houten",  "van houten",  "van houten", "unilever",  "unilever",  "unilever",  "unilever"))
refine1 <- left_join(refine, sort_company_df, by = "company")
refine1_1 <- select(refine1, "fixed_company", "Product code / number", "address", "city", "country", "name")
names(refine1_1)[names(refine1_1) == "fixed_company"] <- "company"
```
``` {R echo = FALSE}
kable(refine1_1)
```

### 2: Separate product code and number

Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

  + Used separate() to separate based on "-"
  
``` {R eval = TRUE}
refine2 <-separate(refine1_1, "Product code / number", c("product_code", "product_number"), sep = "-")
```
``` {R echo = FALSE}
kable(refine2 <-separate(refine1_1, "Product code / number", c("product_code", "product_number"), sep = "-"))
```

### 3: Add product categories

You learn that the product codes actually represent the following product categories:

  + p = Smartphone
  + v = TV
  + x = Laptop
  + q = Tablet

In order to make the data more readable, add a column with the product category for each record.

  + Created new dataframe to have the product code line up with product category
  + Used left_join to find matches

``` {R eval = TRUE}
prod_cod <- data_frame(product_code = c("p", "v", "x", "q"), product_category = c("Smartphone", "TV", "Laptop", "Tablet"))
```
``` {R echo = FALSE}
prod_cod
```
``` {R eval = FALSE}
refine2 <- left_join(refine2, prod_cod, by = "product_code")
```
``` {R echo = FALSE}
kable(refine2 <- left_join(refine2, prod_cod, by = "product_code"))
```

### 4: Add full address for geocoding

You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

  + Used unite() to combine full address

``` {R eval = FALSE}
refine2 <- unite(refine2, "full_address", address, city, country, sep = ",")
```
``` {R echo = FALSE}
kable(refine2 <- unite(refine2, "full_address", address, city, country, sep = ","))
```

### 5: Create dummy variables for company and product category

Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.,

Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever.

  + Created dataframe for company columns with company names as row 1 and corresponding company columns with binary values

``` {R eval = TRUE}
(ccode <- data_frame(company = c("philips", "akzo", "van_houten", "unilever"), product_philips = c(1, 0, 0, 0), product_akzo = c(0, 1, 0, 0), product_van_houten = c(0, 0, 1, 0), product_unilever = c(0, 0, 0, 1)))
refine3 <- left_join(refine2, ccode, by = "company")
```

Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet.

  + Created dataframe for product code columns with product code as row 1 and corresponding product code columns with binary values

``` {R eval = TRUE}
(pcode <- data_frame(product_code = c("p", "v", "x", "q"), product_smartphone = c(1, 0, 0, 0), product_tv = c(0, 1, 0, 0), product_laptop = c(0, 0, 1, 0), product_tablet = c(0, 0, 0, 1)))
refine3 <- left_join(refine3, pcode, by = "product_code")
```

``` {R echo = FALSE}
kable(refine3)
```

### 6: Submit the project on Github

Include your code, the original data as a CSV file refine_original.csv, and the cleaned up data as a CSV file called refine_clean.csv.
